Summary of the things that I learned

### 2024-10-07 100-introduction.tex
Fix title case

The major contributions of this thesis are
\begin{itemize}
  \item Of particular interest is the discussion of Hyland's proof of the locally cartesian closedness of a category of retracts (or Karoubi envelope) in Section \ref{sec:locally-cartesian-closed}, which was proved earlier by Paul Taylor in a different way. It turns out that in the translation from set theory to univalent foundations, the definition of the category of retracts `branches' into two nonequivalent definitions, and that Taylor's proof is about one definition, whereas Hyland's proof is about the other.
  \item \TODO
\end{itemize}

>>>>

The major contributions of this thesis are
\begin{itemize}
  \item Of particular interest is the discussion of Hyland's proof of the relative cartesian closedness of a category of retracts (or Karoubi envelope) in Section \ref{sec:relatively-cartesian-closed}, which was proved earlier by Paul Taylor in a different way. It turns out that in the translation from set theory to univalent foundations, the definition of the category of retracts `branches' into two nonequivalent definitions, and that Taylor's proof is about one definition, whereas Hyland's proof is about the other.
  \item \TODO
\end{itemize}

### 2024-10-07 200-category-theory.tex
Process most of Maarten's feedback
Fix title case

\section{Adjunctions and equivalences}

@@

\begin{lemma}[\coqident{CategoryTheory.YonedaExponentials}{yoneda_preserves_exponentials}]\label{lem:Yoneda-preserves-exponentials}
  The Yoneda embedding preserves exponentials.
\end{lemma}
\begin{proof}
  First of all, note that for $ X, Y, Z : \C $, we have a sequence of isomorphisms  (\autocite{stackexchange:yoneda-exponentials})
  \begin{align*}
    \yo(X^Y)(Z) &\cong P \C(\yo(Z), \yo(X^Y))\\
    &\cong \C(Z, X^Y)\\
    &\cong \C(Z \times Y, X)\\
    &\cong P \C(\yo(Z \times Y), \yo(X))\\
    &\cong P \C(\yo(Z) \times \yo(Y), \yo(X))\\
    &\cong P \C(\yo(Z), \yo(X)^\yo(Y))\\
    &\cong (\yo(X)^\yo(Y))(Z)
  \end{align*}
  using (once or twice) the Yoneda lemma, fully faithfulness of the Yoneda embedding, the property of the exponential object, and the fact that the Yoneda embedding preserves binary products.

@@

\section{(Co)slice categories}
Given an object in a category $ X: \C $, the morphisms to and from $ X $ constitute the slice and coslice categories
\begin{definition}
  The \iindex{slice category} $ \C \downarrow X $ is the category with as objects the morphisms to $ X $:
  \[ (\C \downarrow X)_0 = \sum_{Y: \C} \C(Y, X). \]
  The morphisms from $ (Y_1, f_1) $ to $ (Y_2, f_2) $ are the morphisms $ g: Y_1 \to Y_2 $ making the following diagram commute.
  \begin{center}
    \begin{tikzcd}
      Y_1 \arrow[rr, "g"] \arrow[rd, "f_1"'] & & Y_2 \arrow[ld, "f_2"]\\
      & X
    \end{tikzcd}
  \end{center}
\end{definition}
The \textit{coslice category}\index{slice category!co-} $ X \downarrow \C $ is similar, but with the morphisms \textit{from} $ X $ instead of \textit{to} $ X $:
\[ (X \downarrow \C)_0 = \sum_{Y: \C} \C(X, Y). \]

@@

\section{Dependent products and sums}\label{sec:dependent-products}
The following is based loosely on Section 4.1 of \autocite{taylor}.

@@

\section{(Weakly) terminal objects}
\begin{definition}
  If a category has an object $ I $, such that there is a (not necessarily unique) morphism to it from every other object in the category, $ I $ is said to be a \iindex{weakly terminal object}.
\end{definition}

@@

\section{Monoids as categories}\label{sec:monoid-category}
Take a monoid $ M $.
\begin{definition}[\coqident{CategoryTheory.Categories.MonoidToCategory}{monoid_to_category_ob}]
  We can construct a category $ \C_M $ with $ \C_{M0} = \{ \star \} $, $ \C_M(\star, \star) = M $. The identity morphism on $ \star $ is the identity $ 1: M $. The composition is given by multiplication $ g \cdot_{\C_M} f = f \cdot_M g $.
\end{definition}

@@

\section{The Karoubi envelope}
Let $ \C $ be a category and $ X, Y : \C $ objects. We will denote the type of section-retraction pairs of $ Y $ onto $ X $ with
\[ X \triangleleft Y := \sum_{r : \C(Y, X)} \sum_{s : \C(X, Y)} s \cdot r = \id X. \]
Now, note that for $ (r, s) : X \triangleleft Y $, $ r \cdot s: \C(Y, Y) $ is an idempotent morphism, since $ r \cdot s \cdot r \cdot s = r \cdot s $. We can also wonder whether for an idempotent morphism $ f: \C(X, X) $, we can find some $ Y : \C $ and some $ (r, s) : X \triangleleft Y $ such that $ f = r \cdot s $. If this is the case, we say that the idempotent $ f $ \textit{splits}\index{split idempotent}. If $ f $ does not split, we can wonder whether we can find an embedding $ \iota_\C : \C \hookrightarrow \overline \C $ into some category $ \overline \C $ such that the idempotent $ \iota_\C(f): \overline \C(\iota_\C(X), \iota_\C(X)) $ does split. This is one way to arrive at the \textit{Karoubi envelope}:

>>>>

\section{Adjunctions and Equivalences}

@@

\begin{lemma}[\coqident{CategoryTheory.YonedaExponentials}{yoneda_preserves_exponentials}]\label{lem:Yoneda-preserves-exponentials}
  The Yoneda embedding preserves exponentials.
\end{lemma}
\begin{proof}
  First of all, note that for $ X, Y, Z : \C $, we have a sequence of isomorphisms \autocite{stackexchange:yoneda-exponentials}
  \begin{align*}
    \yo(X^Y)(Z) &\cong P \C(\yo(Z), \yo(X^Y))\\
    &\cong \C(Z, X^Y)\\
    &\cong \C(Z \times Y, X)\\
    &\cong P \C(\yo(Z \times Y), \yo(X))\\
    &\cong P \C(\yo(Z) \times \yo(Y), \yo(X))\\
    &\cong P \C(\yo(Z), \yo(X)^\yo(Y))\\
    &\cong (\yo(X)^\yo(Y))(Z)
  \end{align*}
  using (once or twice) the Yoneda lemma, fully faithfulness of the Yoneda embedding, the property of the exponential object, and the fact that the Yoneda embedding preserves binary products.

@@

\section{(Co)slice Categories}
Given an object in a category $ X: \C $, the morphisms to and from $ X $ constitute the slice and coslice categories
\begin{definition}
  The \iindex{slice category} $ \C \downarrow X $ is the category with as objects the morphisms to $ X $:
  \[ (\C \downarrow X)_0 = \sum_{Y: \C} \C(Y, X). \]
  The morphisms from $ (Y_1, f_1) $ to $ (Y_2, f_2) $ are the morphisms $ g: Y_1 \to Y_2 $ making the following diagram commute.
  \begin{center}
    \begin{tikzcd}
      Y_1 \arrow[rr, "g"] \arrow[rd, "f_1"'] & & Y_2 \arrow[ld, "f_2"]\\
      & X
    \end{tikzcd}
  \end{center}
\end{definition}
The \textit{coslice category}\index{slice category!co-} $ X \downarrow \C $ is similar, but with the morphisms \textit{from} $ X $ instead of \textit{to} $ X $:
\[ (X \downarrow \C)_0 = \sum_{Y: \C} \C(X, Y). \]

@@

\section{Dependent Products and Sums}\label{sec:dependent-products}
The following is based loosely on Section 4.1 of \autocite{taylor}.

@@

\section{(Weakly) Terminal Objects}
\begin{definition}
  If a category has an object $ I $, such that there is a (not necessarily unique) morphism to it from every other object in the category, $ I $ is said to be a \iindex{weakly terminal object}.
\end{definition}

@@

\section{Monoids as Categories}\label{sec:monoid-category}
Take a monoid $ M $.
\begin{definition}[\coqident{CategoryTheory.Categories.MonoidToCategory}{monoid_to_category_ob}]
  We can construct a category $ \C_M $ with $ \C_{M0} = \{ \star \} $, $ \C_M(\star, \star) = M $. The identity morphism on $ \star $ is the identity $ 1: M $. The composition is given by multiplication $ g \cdot_{\C_M} f = f \cdot_M g $.
\end{definition}

@@

\section{The Karoubi Envelope}
Let $ \C $ be a category and $ X, Y : \C $ objects. We will denote the type of section-retraction pairs of $ Y $ onto $ X $ with
\[ X \triangleleft Y := \sum_{r : \C(Y, X)} \sum_{s : \C(X, Y)} s \cdot r = \id X. \]
Now, note that for $ (r, s) : X \triangleleft Y $, $ r \cdot s: \C(Y, Y) $ is an idempotent morphism, since $ r \cdot s \cdot r \cdot s = r \cdot s $. We can also wonder whether for an idempotent morphism $ f: \C(X, X) $, we can find some $ Y : \C $ and some $ (r, s) : X \triangleleft Y $ such that $ f = r \cdot s $. If this is the case, we say that the idempotent $ f $ \textit{splits}\index{split idempotent}. If $ f $ does not split, we can wonder whether we can find an embedding $ \iota_\C : \C \hookrightarrow \overline \C $ into some category $ \overline \C $ such that the idempotent $ \iota_\C(f): \overline \C(\iota_\C(X), \iota_\C(X)) $ does split. This is one way to arrive at the \textit{Karoubi envelope}:

### 2024-10-07 300-univalent-foundations.tex
Fix title case

\section{Dependent type theory}
Univalent foundations takes place in a framework of type theory. In this section, we will quickly introduce some topics that we will need in subsequent sections.

@@

\section{The univalence principle}\label{sec:univalence-principle}
\enquote{\textit{Isomorphic objects are equivalent.}}

@@

\section{The univalence axiom}
Now, even for a basic category, like the category of types, it seems impossible to prove that the univalence principle holds. However, this is no surprise: It turns out that there is a model in set theory for Martin-Löf type theory with univalence \autocite{simplicial-set-model}, but also for Martin-Löf type theory with proof irrelevance \autocite{proof-irrelevant-model}. In the latter, the type $ x = y $ has at most one element for all $ x, y : T $ and all types $ T $, which contradicts univalence. Therefore, univalence is independent of the axioms of Martin-Löf type theory: both univalence or its negation cannot be proved, but either can be assumed as an axiom.

@@

\section{hProps and hSets}\label{sec:homotopy-props-sets}
If we have a type $ T $ and objects $ x $ and $ y $, we can wonder how many elements $ x = y $ has. In set-based mathematics, this would be a nonsensical question: two elements of a set are either equal or not equal. Therefore, we can expect the answer to be that $ x = y $ has at most one element. And indeed, if we do not assume the univalence axiom, we can assume another axiom, called `uniqueness of identity proofs', which states that for $ p, q: x = y $, we have a proof of equality $ h: p = q $.

@@

\section{Truncation and (mere) existence}
As mentioned before, if we want to do mathematics in a dependent type theory, we can `encode' propositions as types. The elements of the type correspond to the proofs that the proposition is true, see for example the identity types. However, we need to be careful here about the distinction between types in general and mere propositions:

@@

\section{Equality and homotopy}
Univalent Foundations is often mentioned together with Homotopy Type Theory, because they are related but distinct concepts. Therefore, we will mention it here.

>>>>

\section{Dependent Type Theory}
Univalent foundations takes place in a framework of type theory. In this section, we will quickly introduce some topics that we will need in subsequent sections.

@@

\section{The Univalence Principle}\label{sec:univalence-principle}
\enquote{\textit{Isomorphic objects are equivalent.}}

@@

\section{The Univalence Axiom}
Now, even for a basic category, like the category of types, it seems impossible to prove that the univalence principle holds. However, this is no surprise: It turns out that there is a model in set theory for Martin-Löf type theory with univalence \autocite{simplicial-set-model}, but also for Martin-Löf type theory with proof irrelevance \autocite{proof-irrelevant-model}. In the latter, the type $ x = y $ has at most one element for all $ x, y : T $ and all types $ T $, which contradicts univalence. Therefore, univalence is independent of the axioms of Martin-Löf type theory: both univalence or its negation cannot be proved, but either can be assumed as an axiom.

@@

\section{Propositions and Sets}\label{sec:homotopy-props-sets}
If we have a type $ T $ and objects $ x $ and $ y $, we can wonder how many elements $ x = y $ has. In set-based mathematics, this would be a nonsensical question: two elements of a set are either equal or not equal. Therefore, we can expect the answer to be that $ x = y $ has at most one element. And indeed, if we do not assume the univalence axiom, we can assume another axiom, called `uniqueness of identity proofs', which states that for $ p, q: x = y $, we have a proof of equality $ h: p = q $.

@@

\section{Truncation and (Mere) Existence}
As mentioned before, if we want to do mathematics in a dependent type theory, we can `encode' propositions as types. The elements of the type correspond to the proofs that the proposition is true, see for example the identity types. However, we need to be careful here about the distinction between types in general and mere propositions:

@@

\section{Equality and Homotopy}
Univalent Foundations is often mentioned together with Homotopy Type Theory, because they are related but distinct concepts. Therefore, we will mention it here.

### 2024-10-07 500-previous-work.tex
Process most of Maarten's feedback
Fix title case

\chapter{Previous work in categorical semantics}\label{ch:previous-work}

@@

\section{The correspondence between categories and typed \texorpdfstring{$ \lambda $}{lambda}-calculi}\label{sec:lambek-correspondence}
In \autocite{curry}, page 413, Scott and Lambek argue that there is a correspondence between simply typed $ \lambda $-calculi and cartesian closed categories (categories with products and `function objects').

@@

\section{The category of retracts}\label{sec:retracts-category}

@@

\section{The \texorpdfstring{$ \lambda $}{lambda}-calculus is algebraic}
As we mentioned in Remark \ref{rem:lambda-theory-or-algebra}, using tools from universal algebra, there are two approaches to study $ \lambda $-calculus-like structures. One can either study $ \lambda $-theories or algebras for the algebraic theory $ \Lambda $. We will see that Hyland shows that these are equivalent.

In his proof, he refers to the 2002 paper `The lambda calculus is algebraic' by Peter Selinger (\autocite{selinger-lambda-calculus-algebraic}). We will not study this paper in detail, but it is interesting to have a quick look at it, because it seems that Hyland took some inspiration from it.

>>>>

\chapter{Previous Work in Categorical Semantics}\label{ch:previous-work}

@@

\section{The Correspondence Between Categories and Typed \texorpdfstring{$ \lambda $}{lambda}-calculi}\label{sec:lambek-correspondence}
In \autocite{curry}, page 413, Scott and Lambek argue that there is a correspondence between simply typed $ \lambda $-calculi and cartesian closed categories (categories with products and `function objects').

@@

\section{The Category of Retracts}\label{sec:retracts-category}

@@

\section{The \texorpdfstring{$ \lambda $}{lambda}-calculus is Algebraic}
As we mentioned in Remark \ref{rem:lambda-theory-or-algebra}, using tools from universal algebra, there are two approaches to study $ \lambda $-calculus-like structures. One can either study $ \lambda $-theories or algebras for the algebraic theory $ \Lambda $. We will see that Hyland shows that these are equivalent.

In his proof, he refers to the 2002 paper `The lambda calculus is algebraic' by Peter Selinger \autocite{selinger-lambda-calculus-algebraic}. We will not study this paper in detail, but it is interesting to have a quick look at it, because it seems that Hyland took some inspiration from it.

### 2024-10-07 600-paper.tex
Fix title case

\chapter{Hyland's paper}\label{ch:the-paper}

@@

\section{Relations between the categories}

@@

\section{Locally cartesian closedness of the category of retracts}\label{sec:locally-cartesian-closed}

@@

\section{An elementary proof of the `Fundamental Theorem'}\label{sec:elementary-fundamental-theorem}
The final theorem that Hyland works towards in his paper constructs an equivalence between two ways to reason about the $ \lambda $-calculus using these tools from universal algebra: $ \lambda $-theories and $ \Lambda $-algebras. In this section, we will provide an elementary proof of this theorem. Then, in the next section, we will give Hyland's original proof, which uses more high-level categorical reasoning.

@@

\section{Hyland's proof}

@@

\section{Theory of extensions}

>>>>

\chapter{Hyland's Paper}\label{ch:the-paper}

@@

\section{Relations Between the Categories}

@@

\section{Relative Cartesian Closedness of the Category of Retracts}\label{sec:relatively-cartesian-closed}

@@

\section{An Elementary Proof of the `Fundamental Theorem'}\label{sec:elementary-fundamental-theorem}

The final theorem that Hyland works towards in his paper constructs an equivalence between two ways to reason about the $ \lambda $-calculus using these tools from universal algebra: $ \lambda $-theories and $ \Lambda $-algebras. In this section, we will provide an elementary proof of this theorem. Then, in the next section, we will give Hyland's original proof, which uses more high-level categorical reasoning.

@@

\section{Hyland's Proof}

@@

\section{The Theory of Extensions}

### 2024-10-07 700-formalization.tex
Fix title case

% 996 0; 447 37; 1383 1017; 26 26; 6313 889; 2562 2332; 1255 326; 3015 1; 4206 1668; 940 48; 9 5; 508 4; 1701 1267; Total: 23361 7620
% 434 292; 1144 58; 96 0; 469 113; 145 42; 495 90; 508 221; Total: 3291 816
\chapter{The formalization}\label{ch:the-formalization}
In addition to the reading, investigating and writing mathematics that lead up to the previous chapters, this thesis project also had a formalization component. Parts of Hyland's paper were carefully written out in detail in a proof assistant and added to a library of formalized mathematics with the univalent point of view, called \textit{UniMath}.

@@

\section{Some numbers}
The code for this project was written over the course of about 18 months, spread over \href{https://github.com/UniMath/UniMath/pulls?q=is%3Apr+author%3Aarnoudvanderleer+is%3Aclosed}{20 pull requests}: 13 with content about Hyland's paper, adding 23.361 and removing 7.620 lines of code, and 7 with only some missing category theoretical preliminaries, adding 3.291 and removing 816 lines.

\section{Overview of the formalized material}
The material that was formalized can be subdivided into material that is introduced or described in Hyland's paper, and category theoretical preliminaries that are necessary to make the proofs in Hyland's paper work.

@@

\section{Displayed categories}\label{sec:displayed-categories}

@@

\section{Duplication in definitions}
One of the adages in software engineering is `DRY': ``Don't Repeat Yourself'': Long expressions and blocks of code that occur multiple times throughout the program should usually be abstracted into a separate function. One of the reasons for this is that it is easier to change code in this function, than to change every instance of the repeated expression or block of code. When writing mathematics, such functions are usually called `lemmas'. However, in this project there was also another example of duplication, which occurred in statements of definitions and lemmas. For example, in the definition of the displayed category of algebraic theories, the displayed object type over a sequence of sets $ T_n $ is
\[ \left(\prod_n \prod_{i : \{ 1, \dots, n \}} T_n \right) \times \left( \prod_{m, n} T_m \times T_n^m \to T_n \right), \]
corresponding to the variables and the substitution. Then, the constructor of an algebraic theory takes arguments
\[ v : \prod_n \prod_{i : \{ 1, \dots, n \}} T_n \quad \text{and} \quad s : \prod_{m, n} T_m \times T_n^m \to T_n. \]
Also, given an algebraic theory, we have accessors:
\[ x_{n, i} : T_n \quad \text{and} \quad \bullet_{m, n} : (\prod_{m, n} T_m \times T_n^m \to T_n) \]
Lastly, when we define a new algebraic theory, we need to provide terms of these types again.

@@

\section{The formalization of the \texorpdfstring{$ \lambda $}{lambda}-calculus} \label{sec:axiomatic-lambda-calculus}
One of the important classes of objects in this thesis is the class of $ \Lambda $-algebras: algebras for the initial $ \lambda $-theory $ \Lambda $, corresponding to the `pure' $ \lambda $-calculus. Therefore, much of this project required the pure $ \lambda $-calculus to be formalized. Now, in this thesis, it is defined as a quotient of an inductive type. However, in the UniMath library, the use of inductive types (other than a handful of basic ones like \texttt{empty}, \texttt{bool}, \texttt{nat} and \texttt{coprod}) is prohibited
\footnote{This is because allowing inductive types requires a much larger trusted codebase of the proof assistant, and because most of mathematics can be formalized with $ \sum $-types or using initial algebras of functors, instead of inductive and record types.}. Also, as we saw in section \ref{sec:quotients}, working with quotients can be somewhat complicated.

@@

\section{The learning curve}
Let me conclude this chapter with a more personal note about formalization. In the past fifteen years, I have worked with a plethora of different programming languages, and I would like to think that in all this time, I have learned a great deal about programming. When I start working with a new language, it takes me a couple of minutes to figure things out, and within an hour, I can probably get a small program up and running, with some help from google. However, when I started working with coq and UniMath, I kind of had to start from scratch again. As it turns out, the naive or direct approach is often not the right one when working with a proof assistant, and of the code that I wrote in the first months, almost nothing remains. Even though every programming language has some sort of learning curve, formalizing in a proof assistant is especially unforgiving: You often have to change an old definition, or start over when you are halfway through a proof, because your current approach gets bogged down further and further with, for example, unintelligible transports. For every line of code that ends up in a pull request, many lines are written and erased again.

>>>>

% 996 0; 447 37; 1383 1017; 26 26; 6313 889; 2562 2332; 1255 326; 3015 1; 4206 1668; 940 48; 9 5; 508 4; 1701 1267; Total: 23361 7620
% 434 292; 1144 58; 96 0; 469 113; 145 42; 495 90; 508 221; Total: 3291 816
\chapter{The Formalization}\label{ch:the-formalization}

In addition to the reading, investigating and writing mathematics that lead up to the previous chapters, this thesis project also had a formalization component. Parts of Hyland's paper were carefully written out in detail in a proof assistant and added to a library of formalized mathematics with the univalent point of view, called \textit{UniMath}.

@@

\section{Some Numbers}
The code for this project was written over the course of about 18 months, spread over \href{https://github.com/UniMath/UniMath/pulls?q=is%3Apr+author%3Aarnoudvanderleer+is%3Aclosed}{20 pull requests}: 13 with content about Hyland's paper, adding 23.361 and removing 7.620 lines of code, and 7 with only some missing category theoretical preliminaries, adding 3.291 and removing 816 lines.

\section{Overview of the Formalized Material}
The material that was formalized can be subdivided into material that is introduced or described in Hyland's paper, and category theoretical preliminaries that are necessary to make the proofs in Hyland's paper work.

@@

\section{Displayed Categories}\label{sec:displayed-categories}

@@

\section{Duplication in Definitions}
One of the adages in software engineering is `DRY': ``Don't Repeat Yourself'': Long expressions and blocks of code that occur multiple times throughout the program should usually be abstracted into a separate function. One of the reasons for this is that it is easier to change code in this function, than to change every instance of the repeated expression or block of code. When writing mathematics, such functions are usually called `lemmas'. However, in this project there was also another example of duplication, which occurred in statements of definitions and lemmas. For example, in the definition of the displayed category of algebraic theories, the displayed object type over a sequence of sets $ T_n $ is
\[ \left(\prod_n \prod_{i : \{ 1, \dots, n \}} T_n \right) \times \left( \prod_{m, n} T_m \times T_n^m \to T_n \right), \]
corresponding to the variables and the substitution. Then, the constructor of an algebraic theory takes arguments
\[ v : \prod_n \prod_{i : \{ 1, \dots, n \}} T_n \quad \text{and} \quad s : \prod_{m, n} T_m \times T_n^m \to T_n. \]
Also, given an algebraic theory, we have accessors:
\[ x_{n, i} : T_n \quad \text{and} \quad \bullet_{m, n} : (\prod_{m, n} T_m \times T_n^m \to T_n) \]
Lastly, when we define a new algebraic theory, we need to provide terms of these types again.

@@

\section{The Formalization of the \texorpdfstring{$ \lambda $}{lambda}-calculus} \label{sec:axiomatic-lambda-calculus}
One of the important classes of objects in this thesis is the class of $ \Lambda $-algebras: algebras for the initial $ \lambda $-theory $ \Lambda $, corresponding to the `pure' $ \lambda $-calculus. Therefore, much of this project required the pure $ \lambda $-calculus to be formalized. Now, in this thesis, it is defined as a quotient of an inductive type. However, in the UniMath library, the use of inductive types (other than a handful of basic ones like \texttt{empty}, \texttt{bool}, \texttt{nat} and \texttt{coprod}) is prohibited
\footnote{This is because allowing inductive types requires a much larger trusted codebase of the proof assistant, and because most of mathematics can be formalized with $ \sum $-types or using initial algebras of functors, instead of inductive and record types.}. Also, as we saw in section \ref{sec:quotients}, working with quotients can be somewhat complicated.

@@

\section{The Learning Curve}
Let me conclude this chapter with a more personal note about formalization. In the past fifteen years, I have worked with a plethora of different programming languages, and I would like to think that in all this time, I have learned a great deal about programming. When I start working with a new language, it takes me a couple of minutes to figure things out, and within an hour, I can probably get a small program up and running, with some help from google. However, when I started working with coq and UniMath, I kind of had to start from scratch again. As it turns out, the naive or direct approach is often not the right one when working with a proof assistant, and of the code that I wrote in the first months, almost nothing remains. Even though every programming language has some sort of learning curve, formalizing in a proof assistant is especially unforgiving: You often have to change an old definition, or start over when you are halfway through a proof, because your current approach gets bogged down further and further with, for example, unintelligible transports. For every line of code that ends up in a pull request, many lines are written and erased again.

### 2024-10-07 910-alternative-definitions.tex
Fix title case

\section{Lawvere theory}
\begin{definition}
  An \iindex{algebraic theory} as presented in \autocite{algebraic-theories-2010} is a small category with finite products.
\end{definition}
\begin{definition}
  An \iindex{algebra} for an algebraic theory $ T $ is a finite-product-preserving functor $ T \to \SET $.
\end{definition}

@@

\section{Monoid in a skew-monoidal category}
For details and a general treatment, see \autocite{monads-endofunctors}, Section 3 and specifically Theorem 3.4.

>>>>

\section{Lawvere Theory}
\begin{definition}
  An \iindex{algebraic theory} as presented in \autocite{algebraic-theories-2010} is a small category with finite products.
\end{definition}
\begin{definition}
  An \iindex{algebra} for an algebraic theory $ T $ is a finite-product-preserving functor $ T \to \SET $.
\end{definition}

@@

\section{Monoid in a Skew-Monoidal Category}
For details and a general treatment, see \autocite{monads-endofunctors}, Section 3 and specifically Theorem 3.4.
